---
title: "Cardiovascular Disease"
author: "Diana, Daria, Luna Nguyen"
date: "11/14/2022"
output: html_document
---

```{r source, echo= FALSE, message = FALSE}
source("../source/summary_info.R")
```

### Washington Cardiovascular Disease Death Statistics

For our project, we investigated a data set about the rates and trends of cardiovascular disease in the
United States in search of possible risk factors or other power issues. We narrowed our data down to
Washington state and discovered that the county with the highest deaths of 65+ year olds per 100,000
people was `r summary_info$county_most_65_above_deaths_per_100000_in_wa$LocationDesc` in 
`r summary_info$county_most_65_above_deaths_per_100000_in_wa$Year` with 
`r round(summary_info$county_most_65_above_deaths_per_100000_in_wa$Data_Value)` 
deaths, while the county with the lowest number of deaths per 100,000 people was 
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$LocationDesc` in 
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$Year` with 
`r round(summary_info$county_least_65_above_deaths_per_100000_in_wa$Data_Value)` 
deaths. Additionally, the county with the least deaths of 35-64 year olds per 100,000 was 
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$LocationDesc` in 
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Year` with 
`r round(summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Data_Value)` 
deaths, while the county with the lowest number of deaths of 35-64 year olds per 100,000 people was 
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$LocationDesc` in 
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Year` with 
`r round(summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Data_Value)` 
deaths. We also found that the county with the highest percent change in death rate in 2010-2019 was 
`r summary_info$county_most_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of 
`r round(summary_info$county_most_percent_death_rate_chge_10_to_19$Data_Value)` and the county with
the lowest percentage change in death rate in 2010-2019 was 
`r summary_info$county_least_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of 
`r round(summary_info$county_least_percent_death_rate_chge_10_to_19$Data_Value)`.

### Overall Changes in Heart Disease and Stroke Death From 2000-2019

```{r table, echo=FALSE}
library("knitr")
source("../source/table.R")
kable(wa_table)
```

I included a table of the overall changes in heart disease and stroke death from 2000-2019 because it can
be helpful to analyze the changes in stroke and heart disease deaths over a period of time. It reveals
that Clallam county has the biggest increase in those deaths while Douglas has the biggest decrease in
deaths. Also, overall, most counties in Washington have had an increase in stroke and heart disease
deaths, which will be helpful with our research question about how has cardiovascular disease and stroke
fatalities changed overtime.

### Graphs About the Deaths Rates of 65+ Year Olds in Washington 2019

```{r map, echo=FALSE}
source("../source/map.R")
wa_2019_plot
```

This map reveals the rates at which people die from heart disease and stroke per 100,000 people in each
county. I included the chart because information about what counties have the highest and lowest rates
are be easily seen with the bluest counties being the highest and the darkest being the lowest. It
reveals counties where rates are highest and lowest. The southern counties seem to have the higher rates
with Lewis county being the highest while the northern and northeastern counties have the lower rates.
This data will be helpful in answering one of our research questions about why there are differences in
rates, specifically if income is a factor.

```{r pie chart, echo=FALSE}
source("../source/piechart.R")
piechart
```

This is a pie chart that shows the number of deaths per 100,000 people for people ages 65+ in Washington
during 2019 and a percentage representing how many of those people were male or female. From the pie
chart, we can see that for every 100,000 people, slightly more than 30,000 people ages 65+ die and among
these roughly 30,000 people, 53% of them are male and 47% are female. This chart is included because our
project seeks to find possible risk factors for cardiovascular disease and we wanted to test if there was
more instances of disease in people of either sex. As you can see from the chart, we found that a
possible risk factor for death by cardiovascular disease in Washington is being male. By identifying this
information, we can further investigate data revolving around the sex of people who have died of
cardiovascular disease and determine if the disparity was due to chance or other underlying factors.

```{r bar plot, echo= FALSE}
source("../source/group-project-bar-graph.R")
bargraph
```

This is a bar chart that shows how many deaths there are per 100,000 people ages 65+ in Washington during
2019 for different races. From this data set we can see that white people seem to die the most out of
every other race by an extremely high amount, Hispanics die the second most black people and American
Indians die the least.However we suspect that it's because white people heart diseases reporting more
because they typically go to a hospital versus other races will die home because I don't have the money
to go to hospital. We can see by looking at the original unfiltered data set that a lot of the data on
other races other than white are non-existent.
