---
title: "Cardiovascular Disease"
author: "Diana, Daria, Luna Nguyen"
date: "11/14/2022"
output: html_document
---

```{r source, echo= FALSE, message = FALSE}
source("../source/summary_info.R")
```

### Washington Cardiovascular Disease Death Statistics

For our project, we investigated a data set about the rates and trends of cardiovascular disease in the
United States in search of possible risk factors or other power issues. We narrowed our data down to
Washington state and discovered that the county with the highest deaths of 65+ year olds per 100,000
people was `r summary_info$county_most_65_above_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_most_65_above_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_most_65_above_deaths_per_100000_in_wa$Data_Value)`
deaths, while the county with the lowest number of deaths per 100,000 people was
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_least_65_above_deaths_per_100000_in_wa$Data_Value)`
deaths. Additionally, the county with the least deaths of 35-64 year olds per 100,000 was
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Data_Value)`
deaths, while the county with the lowest number of deaths of 35-64 year olds per 100,000 people was
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Data_Value)`
deaths. We also found that the county with the highest percent change in death rate in 2010-2019 was
`r summary_info$county_most_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of
`r round(summary_info$county_most_percent_death_rate_chge_10_to_19$Data_Value)` and the county with
the lowest percentage change in death rate in 2010-2019 was
`r summary_info$county_least_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of
`r round(summary_info$county_least_percent_death_rate_chge_10_to_19$Data_Value)`.

### Overall Changes in Heart Disease and Stroke Death From 2000-2019

```{r table, echo=FALSE}
library("knitr")
source("../source/table.R")
kable(wa_table)
```

I included a table of the overall changes in heart disease and stroke death from 2000-2019 because it can
be helpful to analyze the changes in stroke and heart disease deaths over a period of time. It reveals
that Clallam county has the biggest increase in those deaths while Douglas has the biggest decrease in
deaths. Also, overall, most counties in Washington have had an increase in stroke and heart disease
deaths, which will be helpful with our research question about how has cardiovascular disease and stroke
fatalities changed overtime.

### Graphs About the Deaths Rates of 65+ Year Olds in Washington 2019

```{r map, echo=FALSE}
source("../source/map.R")
wa_2019_plot
```

This map reveals the rates at which people die from heart disease and stroke per 100,000 people in each
county. I included the chart because information about what counties have the highest and lowest rates
are be easily seen with the bluest counties being the highest and the darkest being the lowest. It
reveals counties where rates are highest and lowest. The southern counties seem to have the higher rates
with Lewis county being the highest while the northern and northeastern counties have the lower rates.
This data will be helpful in answering one of our research questions about why there are differences in
rates, specifically if income is a factor.

```{r pie chart, echo=FALSE}
source("../source/scatterplot.R")
scatterplot
```

This is a scatter plot that shows the median deaths per 100,000 people for people ages 65+ in Washington
over the span of 2000 to 2019. This scatter plot was created in hopes that it would reveal any
significant changes in the median rates of death due to cardiovascular disease. We could then
investigate these time ranges for more information about any potential risk factors or events that caused
the significant decrease or increase in the median death rates. From the scatter plot we can see that in
2008 Washington had the highest median death rate. The years leading up to 2008 also have considerably
high numbers compared to 2000. This scatter plot allows us to see that something happened between the
years 2000 and 2008 that caused a large increase. Additionally, we can see that after 2008, the median
death rates slowly starting dropping below 400 before rising above again in 2017. Both these time ranges 
should be further investigated because there may be underlying events that caused these fluctuations,
like how Washington State implemented a sugar tax.

```{r bar plot, echo= FALSE, message = FALSE}
source("../source/group-project-bar-graph.R")
bargraph
```

This is a bar graph that shows the median number of deaths of 65+ year olds per 100,000 people in Washington during the year 2019. We created this bar graph because we wanted to assess if race had any influence on the risk on dying to cardiovascular
disease. Our bar graph shows Black people are the number one race that died from cardiovascular disease. The graph shows that they die SIGNIFICANTLY more than any other race. However, this may be due to there being a small amount of observations on Black people in our data set. For the 2nd most deaths, White people and American Indian and Alaska Native are a close tie but White people are a little higher. For the races with the least cardiovascular disease death Hispanic and Asian and Pacific Islander are a close tie with Asian and Pacific Islander being slightly higher. 


Analysis:
American Indian and Alaska Natives having the same death
rates as White people is suspicious because other research has shown that American Indian and Alaskan Natives live in poor reservations. After investigating our data set we discovered that, like Black people, American Indians and Alaskan Natives do not have many observations. This is probably due to there not being a high American Indian and Alaskan Native population due to the genocide that happened before this data was collected and therefore does not have much data on American Indians. We can see that original dataset had a lot of NAs with the American Indian population. Our data set has the most data on the white population. Their high population is going to make their deaths go up but white privilege is going to make their deaths go down. Our data set consists mostly of data on the white population.


Perhaps if we took data from all over the world instead of just
Washington, we would have reached a different conclusion. However, our graph does reveal to us that there seems to be a possible risk factor for Black people and a slight disparity in the death rates of White people that should be investigated further.

*The correct term for American Indian is indigenous people