---
title: "Cardiovascular Disease"
author: "Diana, Daria, Luna Nguyen"
date: "11/14/2022"
output: html_document
---

```{r source, echo= FALSE, message = FALSE}
source("../source/summary_info.R")
```

### Washington Cardiovascular Disease Death Statistics

For our project, we investigated a data set about the rates and trends of cardiovascular disease in the
United States in search of possible risk factors or other power issues. We narrowed our data down to
Washington state and discovered that the county with the highest deaths of 65+ year olds per 100,000
people was `r summary_info$county_most_65_above_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_most_65_above_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_most_65_above_deaths_per_100000_in_wa$Data_Value)`
deaths, while the county with the lowest number of deaths per 100,000 people was
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_least_65_above_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_least_65_above_deaths_per_100000_in_wa$Data_Value)`
deaths. Additionally, the county with the least deaths of 35-64 year olds per 100,000 was
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_most_35_to_64_deaths_per_100000_in_wa$Data_Value)`
deaths, while the county with the lowest number of deaths of 35-64 year olds per 100,000 people was
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$LocationDesc` in
`r summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Year` with
`r round(summary_info$county_least_35_to_64_deaths_per_100000_in_wa$Data_Value)`
deaths. We also found that the county with the highest percent change in death rate in 2010-2019 was
`r summary_info$county_most_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of
`r round(summary_info$county_most_percent_death_rate_chge_10_to_19$Data_Value)` and the county with
the lowest percentage change in death rate in 2010-2019 was
`r summary_info$county_least_percent_death_rate_chge_10_to_19$LocationDesc` with a percentage of
`r round(summary_info$county_least_percent_death_rate_chge_10_to_19$Data_Value)`.

### Overall Changes in Heart Disease and Stroke Death From 2000-2019

```{r table, echo=FALSE}
library("knitr")
source("../source/table.R")
kable(wa_table)
```

I included a table of the overall changes in heart disease and stroke death from 2000-2019 because it can
be helpful to analyze the changes in stroke and heart disease deaths over a period of time. It reveals
that Clallam county has the biggest increase in those deaths while Douglas has the biggest decrease in
deaths. Also, overall, most counties in Washington have had an increase in stroke and heart disease
deaths, which will be helpful with our research question about how has cardiovascular disease and stroke
fatalities changed overtime.

### Graphs About the Deaths Rates of 65+ Year Olds in Washington 2019

```{r map, echo=FALSE}
source("../source/map.R")
wa_2019_plot
```

This map reveals the rates at which people die from heart disease and stroke per 100,000 people in each
county. I included the chart because information about what counties have the highest and lowest rates
are be easily seen with the bluest counties being the highest and the darkest being the lowest. It
reveals counties where rates are highest and lowest. The southern counties seem to have the higher rates
with Lewis county being the highest while the northern and northeastern counties have the lower rates.
This data will be helpful in answering one of our research questions about why there are differences in
rates, specifically if income is a factor.

```{r pie chart, echo=FALSE}
source("../source/scatterplot.R")
scatterplot
```

This is a scatter plot that shows the median deaths per 100,000 people for people ages 65+ in Washington
over the span of 2000 to 2019. This scatter plot was created in hopes that it would reveal any
significant changes in the median rates of death due to cardiovascular disease. We could then
investigate these time ranges for more information about any potential risk factors or events that caused
the significant decrease or increase in the median death rates. From the scatter plot we can see that in
2008 Washington had the highest median death rate. The years leading up to 2008 also have considerably
high numbers compared to 2000. This scatter plot allows us to see that something happened between the
years 2000 and 2008 that caused a large increase. Additionally, we can see that after 2008, the median
death rates slowly starting dropping below 400 before rising above again in 2017. Both these time ranges 
should be further investigated because there may be underlying events that caused these fluctuations,
like how Washington State implemented a sugar tax.

```{r bar plot, echo= FALSE, message = FALSE}
source("../source/group-project-bar-graph.R")
bargraph
```

This is a bar chart that shows how many deaths there are per 100,000 people ages 65+ in Washington during
2019 for different races. From this data set we can see that white people seem to die the most out of
every other race by an extremely high amount, Hispanics die the second most, black people third most and American
Indians die the least.However we suspect that it's because white people heart diseases reporting more
because they typically go to a hospital versus other races will die home because I don't have the money
to go to hospital. We can see by looking at the original unfiltered data set that a lot of the data on
other races other than white are non-existent.
